\section*{Global Position Projection for Constrained Systems}

To achieve a "one-shot" stable stack (replacing Baumgarte stabilization), we formulate the position correction as a global linear system for each independent island.

\subsection*{1. Linearized Constraint Equation}
Let $\Phi(q) = 0$ represent the position-level constraint (e.g., non-penetration). If $\Phi(q) \neq 0$ due to numerical drift, we seek a displacement $\Delta q$ such that:
$$ \Phi(q + \Delta q) = 0 $$
Using a first-order Taylor expansion:
$$ \Phi(q) + J \Delta q \approx 0 \implies J \Delta q = -\Phi(q) $$
where $J = \frac{\partial \Phi}{\partial q}$ is the Jacobian matrix mapping configuration space to constraint space.

\subsection*{2. Principle of Least Constraint}
The correction $\Delta q$ must be consistent with the system's mass distribution. We express the displacement as a function of the Lagrange multipliers $\lambda$ (geometric impulses):
$$ \Delta q = M^{-1} J^T \lambda $$
where $M^{-1}$ is the inverse mass and inertia matrix.

\subsection*{3. The System Matrix (Schur Complement)}
Substituting the displacement into the linearized constraint equation yields the system of equations for the island:
$$ J (M^{-1} J^T \lambda) = -\Phi(q) $$
Let $A = J M^{-1} J^T$ be the symmetric, positive-definite (SPD) system matrix. The final system to solve for $\lambda$ is:
$$ A \lambda = -\Phi(q) $$

\subsection*{4. Sparsity and Parallelization}
\begin{itemize}
    \item \textbf{Pairwise Sparsity:} Since each constraint connects exactly two bodies, each row of $J$ contains only two non-zero blocks. 
    \item \textbf{Structure:} The resulting matrix $A$ is sparse, with non-zero entries existing only where bodies share a common constraint.
    \item \textbf{Solving:} Use a Sparse $LDL^T$ or Cholesky factorization per island for "infinite" stiffness.
\end{itemize}

\subsection*{5. State Synchronization}
Once $\lambda$ is found, the final state is updated to ensure the velocity matches the corrected geometry.

\subsubsection*{5a. Position Update}
Apply position corrections after the integration stage (not before), writing directly to the integrated body state:
$$ q_{new} = q_{integrated} + \Delta q, \qquad \Delta q = M^{-1} J^T \lambda $$
Applying corrections \textit{before} integration causes double-application: the integrate stage computes $q = q_{old} + v \cdot \Delta t$, so any correction folded into $v$ before integration gets integrated into position a second time, producing exponential energy growth.

\subsubsection*{5b. Velocity Synchronization (Normal-Component Replacement)}
The naive formulation $v_{new} = \Delta q / \Delta t$ replaces the \textit{entire} velocity, which destroys tangential velocity computed by the iterative solver (TGS). For contact constraints, the correction $\Delta q$ lies entirely in the constraint-normal subspace (since only normal rows participate in position projection), but the body's velocity also contains tangential components from friction solving that must be preserved.

The correct procedure for each body $k$ in the island:
\begin{enumerate}
    \item Compute the correction velocity: $v_{corr,k} = \Delta q_k / \Delta t$
    \item Gather all contact normals $\{n_i\}$ from constraints touching body $k$.
    \item For each unique contact normal $n_i$:
    \begin{enumerate}
        \item Remove the old normal component: $v_k \leftarrow v_k - (v_k \cdot n_i)\, n_i$
        \item Add the correction's normal component: $v_k \leftarrow v_k + (v_{corr,k} \cdot n_i)\, n_i$
    \end{enumerate}
\end{enumerate}
This projects out the pre-existing constraint-normal velocity and replaces it with the velocity implied by the position correction, while leaving the tangential (friction) velocity intact. The result is a velocity field that is consistent with the corrected geometry without injecting energy or destroying solver work.

\subsubsection*{5c. Why Not Simply Add?}
Adding $v_{corr}$ to the existing velocity would double-count the normal impulse already applied by the iterative solver during TGS. The solver already produced a velocity that \textit{attempts} to resolve penetration; position projection then fixes the residual. Replacing the normal component ensures only the corrected residual is reflected in velocity, avoiding energy injection that would destabilize stacks.

